---
title: "Visualization"
output: html_document
---

## Loading data

```{r}
data <- read.csv("WHRData2021.csv")
library(ggplot2)
require(scales)
```

Full column names:

1. Country name
2. Population 2020
3. Population 2019
4. "COVID-19 deaths per 100,000 population in 2020"
5. Median age
6. Island
7. Index of exposure to COVID-19  infections in other countries as of March 31
8. Log of average distance to SARS countries
9. WHO Western Pacific Region
10. Female head of government
11. Index of institutional trust
12. Gini coefficient of income
13. "All-cause death count, 2017"
14. "All-cause death count, 2018",
15. "All-cause death count, 2019"
16. "All-cause death count, 2020"
17. "Excess deaths in 2020 per 100,000 population, relative to 2017-2019 average"
18. Regional indicator
19. Ladder score
20. Standard error of ladder score
21. upperwhisker
22. lowerwhisker
23. Logged GDP per capita
24. Social support
25. Healthy life expectancy
26. Freedom to make life choices
27. Generosity
28. Perceptions of corruption
29. Ladder score in Dystopia
30. Explained by: Log GDP per capita
31. Explained by: Social support
32. Explained by: Healthy life expectancy
33. Explained by: Freedom to make life choices
34. Explained by: Generosity
35. Explained by: Perceptions of corruption
36. Dystopia + residual

Remove NAs from data (observations that have NAs for healthy life expectancy)
```{r}
remove <- which(is.na(data$Healthy.life.expectancy))
data <- data[-remove, ]
```

```{r}
sort(data$Healthy.life.expectancy, decreasing = TRUE)

for (i in 1:12){
  vector <- c(76.953, 76.820, 75.100, 74.700, 74.400, 74.000, 73.900, 73.900, 73.898, 73.898, 73.800, 73.800)
  print(which (data$Healthy.life.expectancy == vector[i]))
}
```

## Country name (x), Healthy life expectancy (y), Island (color)

Top 12 countries in terms of healthy life expectancy (top 10 scores): 120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62

```{r}
countries <- factor(data$Country.name[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)], levels = data$Country.name[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)], ordered = TRUE)
healthy_life_expect <- data$Healthy.life.expectancy[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)]
island <- factor(data$Island[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)])
island <- ifelse(island == 0, "No", "Yes")

ggplot(
  data = data.frame(data[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62), ]), 
  aes(x = countries, 
      y = healthy_life_expect,
      fill = island,
      group = island
    ) 
) + geom_bar(stat = "identity") +
  scale_y_continuous(limits = c(0, 80)) +
  theme(axis.text.x = element_text(angle = 20)) +
  labs(
    x = "Countries",
    y = "Healthy life expectancy score",
    title = "Countries With the Top 10 Healthy Life Expectancy Scores by Island Status"
     ) + 
   guides(fill = guide_legend(title = "Island?")) +
  scale_fill_manual(values = c("yellowgreen", "deepskyblue2")) 
```

## Country name (x), Healthy life expectancy (y), Female head of government (color)

```{r}
countries <- factor(data$Country.name[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)], levels = data$Country.name[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)], ordered = TRUE)
healthy_life_expect <- data$Healthy.life.expectancy[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)]
female <- factor(data$Female.head.of.government[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)])
female <- ifelse(female == 0, "No", "Yes")

ggplot(
  data = data.frame(data[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62), ]), 
  aes(x = countries, 
      y = healthy_life_expect,
      fill = female,
      group = female
    ) 
) + geom_bar(stat = "identity") + coord_flip() +
  theme(axis.text.x = element_text(angle = 20)) +
  labs(
    x = "Countries",
    y = "Healthy life expectancy score",
    title = "Countries With the Top 10 Healthy Life Expectancy Scores by Female Leader Status"
     ) +
  guides(fill = guide_legend(title = "Female leader?")) +
  scale_fill_manual(values = c("steelblue1", "violetred1")) +
  coord_cartesian(ylim = c(70, 77))
```

## Country name (x), Healthy life expectancy (y), Median age (color)

```{r}
countries <- factor(data$Country.name[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)], levels = data$Country.name[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)], ordered = TRUE)
healthy_life_expect <- data$Healthy.life.expectancy[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)]
median_age <- data$Median.age[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)]

ggplot(
  data = data.frame(data[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62), ]), 
  aes(x = countries, 
      y = healthy_life_expect,
      fill = median_age,
      group = median_age
    ) 
) + geom_bar(stat = "identity") +
  scale_y_continuous(limits = c(0, 80)) +
  theme(axis.text.x = element_text(angle = 20)) +
  labs(
    x = "Countries",
    y = "Healthy life expectancy score",
    title = "Countries With the Top 10 Healthy Life Expectancy Scores by Median Age"
     ) +
  guides(fill = guide_legend(title = "Median age")) + 
  scale_fill_continuous(high = "#132B43", low = "#56B1F7")
```

```{r}
countries <- factor(data$Country.name[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)], levels = data$Country.name[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)], ordered = TRUE)
healthy_life_expect <- data$Healthy.life.expectancy[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)]
median_age <- data$Median.age[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62)]

ggplot(
  data = data.frame(data[c(120, 53, 65, 125, 129, 42, 6, 124, 32, 102, 23, 62), ]), 
  aes(x = median_age, 
      y = healthy_life_expect,
      color = countries,
      group = countries
    ) 
) + geom_point()
  theme(axis.text.x = element_text(angle = 20)) +
  labs(
    x = "Countries",
    y = "Healthy life expectancy score",
    title = "Countries With the Top 10 Healthy Life Expectancy Scores by Median Age"
     ) +
  guides(fill = guide_legend(title = "Median age"))
```

```{r}
regions <- factor(data$Regional.indicator, levels = c("Middle East and North Africa", "Sub-Saharan Africa", "Latin America and Caribbean", "East Asia", "South Asia", "Southeast Asia", "North America and ANZ", "Central and Eastern Europe", "Western Europe", "Commonwealth of Independent States"))

ggplot(
  data = data.frame(data),
  aes(x = Median.age,
      y = Healthy.life.expectancy, 
      color = regions
      )
) + geom_point(size = 3) +
  labs(
    x = "Median age",
    y = "Healthy life expectancy score",
    title = "Healthy Life Expectancy Score vs. Median Age by Region"
     ) + 
  labs(color = "Geographic region") 
```

## Adding the coordinates of countries to the dataset

```{r}
coordinates <- read.csv("world_country_and_usa_states_latitude_and_longitude_values.csv")
coordinates <- data.frame(country = coordinates$country, latitude = coordinates$latitude, longitude = coordinates$longitude)

indices <- coordinates$country %in% data$Country.name
coordinates <- coordinates[indices, ]

new_data <- merge(x = data, y = coordinates, by.x = "Country.name", by.y = "country", all = TRUE)

missing <- which(is.na(new_data$latitude)) # Manually enter the coordinates for these countries
new_data$Country.name[missing]

new_data$latitude[29] <- 21.7587
new_data$longitude[29] <- 4.0383

new_data$latitude[53] <- 114.1694
new_data$longitude[53] <- 22.3193

new_data$latitude[63] <- 5.5471
new_data$longitude[63] <- 7.5400

new_data$latitude[94] <- 95.9560
new_data$longitude[94] <- 21.9162

new_data$latitude[102] <- 33.6577
new_data$longitude[102] <- 35.2480

new_data$latitude[103] <- 21.7453
new_data$longitude[103] <- 41.6086

new_data$latitude[130] <- 120.9605
new_data$longitude[130] <- 23.6978

write.csv(new_data, "new_data.csv", row.names = FALSE)
```

## Adding a column for latlong format (latitude, longitude)
```{r}
new_data$lat_long <- numeric(nrow(new_data))

for (i in 1:nrow(new_data)){
  new_data$lat_long[i] <- paste0(new_data$latitude[i], ",", new_data$longitude[i])
}

write.csv(new_data, "new_data.csv", row.names = FALSE)
```

## Trying to make maps in R instead

`install.packages(c("cowplot", "googleway", "ggplot2", "ggrepel", "ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata"))`

```{r}
library(Rcpp)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) + geom_sf()
```

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
require(maps)
require(viridis)
theme_set(
  theme_void()
  )

new_data <- read.csv("new_data.csv")
world_map <- map_data("world")
ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", colour = "white")

merged_map_data <- merge(x = new_data, y = world_map, by.x = "Country.name", by.y = "region", all = TRUE)

merged_map_data <- data.frame(COVID.19.deaths.per.100.000.population.in.2020 = merged_map_data$COVID.19.deaths.per.100.000.population.in.2020, lat = merged_map_data$lat, long = merged_map_data$long, group = merged_map_data$group, order = merged_map_data$order, subregion = merged_map_data$subregion, Country.name = merged_map_data$Country.name)

#ggplot(merged_map_data, aes(long, lat, group = group)) + geom_polygon(aes(fill = COVID.19.deaths.per.100.000.population.in.2020), color = "white") + scale_fill_viridis_c(option = "C")
```

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
require(maps)
require(viridis)

world <- map_data("world")
new_data <- read.csv("new_data.csv")
names(new_data)[names(new_data) == "COVID.19.deaths.per.100.000.population.in.2020"] <- "COVID-19 deaths 2020"

ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "white", fill = "lightgray", size = 0.1
  ) +
  geom_point(
    data = new_data,
    aes(longitude, latitude, size = `COVID-19 deaths 2020`, color = `COVID-19 deaths 2020`), alpha = 0.7
  ) + scale_colour_gradient(name = "COVID-19 deaths 2020", guide = guide_colourbar(reverse = TRUE), low = "#56B1F7", high = "#132B43") +
  scale_size_continuous(name = "COVID-19 deaths 2020")
```